---
title: "Reproducibility research part 1"
author: "Ymke de Jong"
date: "February 8, 2017"
output: html_document
---
### 1. Code for reading in the dataset and/or processing the data
```{r setup, include=FALSE}
rm(list=ls())
setwd("C:/Users/Ymke.de.jong/OneDrive - BPSolutions Group/Ymke/Scirpts/R/Coursera - Reproducibility Research")

dataACT <- read.csv("activity.csv",stringsAsFactors=FALSE);
dataACT$date <- as.POSIXct(dataACT$date, format="%Y-%m-%d")
dataACT$day <- weekdays(as.Date(dataACT$date))

dataACT_df <- data.frame(dataACT)

rm(dataACT)

head(dataACT_df)
```

### 2. Histogram of the total number of steps taken each day

```{r,fig.width=7}
subgr_data <- aggregate(dataACT_df$steps,list(dataACT_df$date),sum)
names(subgr_data) <- c("date","total")
head(subgr_data)
hist(subgr_data$total, 
     breaks=seq(from=0, to=25000, by=2500))

```


### 3. Mean and median number of steps taken each day

```{r,echo = FALSE}
mn      <- mean(subgr_data$total,na.rm = TRUE)
med     <- median(subgr_data$total,na.rm = TRUE)
```
*The mean of the steps per day is: `r mn ` and the median is `r med`*

### 4. Time series plot of the average number of steps taken

```{r}
subgr_mean <- aggregate(dataACT_df$steps,list(dataACT_df$interval),mean,na.rm=TRUE)
names(subgr_mean) <- c("interval","average")

plot(x=subgr_mean$interval,
     y=subgr_mean$average,
     type = "l",
     lwd = 2,
     xlab = "Interval [minutes]",
     ylab = "Mean steps")
```

### 5. The 5-minute interval that, on average, contains the maximum number of steps

```{r}
posmax <- subgr_mean[which.max(subgr_mean$average),]$interval
```

The interval with the maximum mean steps is `r posmax`.

### 6. Code to describe and show a strategy for imputing missing data

```{r}
N_na <- sum(is.na(dataACT_df$steps))

na_pos <- which(is.na(dataACT_df$steps))

mean_vec <- rep(mean(dataACT_df$steps, na.rm=TRUE), times=length(na_pos))

dataACT_df[na_pos, "steps"] <- mean_vec
```

### 7. Histogram of the total number of steps taken each day after missing values are imputed

```{r,fig.width=7}
subgr_sum <- aggregate(dataACT_df$steps, list(dataACT_df$date),sum)
names(subgr_sum) <- c("date","sum")

hist(subgr_sum$sum,
     breaks=seq(from=0, to=25000, by=2500),
     xlab="Number of steps")
```

### 8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r}
dataACT_df <- cbind(dataACT_df,daytype=ifelse(dataACT_df$day ==
                                                              "Saturday"|
                                        dataACT_df$day == "Sunday", "weekend",
                                        "weekday"))
data_weekdays <- aggregate(dataACT_df$steps,
                           list(dataACT_df$daytype,
                                dataACT_df$day,
                                dataACT_df$interval),
                           mean)

names(data_weekdays) <- c("daytype","day","interval","average")

library(ggplot2)

ggplot(data_weekdays, aes(x=interval, y=average)) + 
        geom_line(color="violet") + 
        facet_wrap(~ daytype, nrow=2, ncol=1) +
        labs(x="Interval", y="Number of steps") 


```
### 9. All of the R code needed to reproduce the results (numbers, plots, etc.) in the report
